---
title: "STAT 385: Homework 03"
author: "Ha Khanh Nguyen - Spring 2020, UIUC"
date: 'Due: Friday, February 21, 11:59 PM'
output:
  prettydoc::html_pretty:
    theme: architect
    toc: true
    
---

## Instructions

- Maximum total number of points for this homework is 35.
- Download the [R script template](hw03-your-name.R) and write your R code in this file for submission. h
- Rename the R script as `hw03-your-name.R`. For example, my R script file will be `hw03-ha-nguyen.R`.
- Make sure to include your name in the R script file.
- Submit the R script (`.R`) file to Compass by **11:59 PM on Friday, 02/21/2020**. You can submit as many times as needed up until the deadline. **The latest version submitted will be graded**.

---

## Exercise 1: Vectorized Code

For each part of this exercise, a function has been written for some specific purposes using loop. Your tasks are:
  - To read/run the original function to understand what it does (the easiest way to do this is to run the function through an example by hand),
  - Then write the **vectorize** the function.

- **(a)** (5 points) Name your vectorized function `specific_sum_vec()`.

```{r}
specific_sum_loop <- function(m, pos) {
  my_sum <- 0
  for (i in 1:nrow(m)) {
    for (j in pos) {
      my_sum <- my_sum + m[i, j]
    }
  }
  my_sum
}
```

Running the function:

```{r}
m <- matrix(data = c(114:133), nrow = 4, byrow = TRUE)
pos <- c(1, 4, 2)
specific_sum_loop(m, pos)
```

```{r eval = FALSE}
# write your specific_sum_vec() function
specific_sum_vec <- function(m,pos) {
  sum_col <- sum(m[,pos])
  sum_col
}
m <- matrix(data = c(114:133), nrow = 4, byrow = TRUE)
pos <- c(1,4,2)
specific_sum_vec(m,pos)
# test your code
```

- **(b)** (5 points) Name your vectorized function `long_name_vec()`.

```{r}
long_names_vec <- function(names) {
  output <- c()
  for (i in 1:length(names)) {
    name <- names[i]
    if (nchar(name) >= 8) {
      output <- c(output, name)
    }
  }
  output
}
```

Running the function:

```{r}
my_names <- c("alexander", "david", "sebastian", "johnathan", "christopher", "ha",
              "washington", "lincoln", "maximo", "mason", "luca", "anthony", "kevin")

long_names_vec(names = my_names)
```

```{r eval = FALSE}
# write your specific_sum_vec() function
long_name_vec <- function(names) {
  names[nchar(names) >= 6]
  names
}

# test your code
my_names <- c("alexander", "david", "sebastian", "johnathan", "christopher", "ha",
              "washington", "lincoln", "maximo", "mason", "luca", "anthony", "kevin")

long_name_vec(names = my_names)
```


- **(c)** (10 points) Name your vectorized function `roulette_vec()`.

```{r}
roulette_loop <- function(many_bets) {
  
  # randomly decide whether you lose or win on each bet
  win_lose_random <- sample(x = c(TRUE, FALSE), size = length(many_bets),
                            replace = TRUE)
  
  total_prize <- c()
  
  # for each bet
  for (i in 1:length(many_bets)) {
    bet <- many_bets[i]
    prize <- 0
    
    # if we win, then check how much the prize is
    if (win_lose_random[i]) {
      if (bet == 'low') {
        prize <- 10
      } else if (bet == 'high'){
        prize <- 10
      } else if (bet == 'red') {
        prize <- 20
      } else if (bet == "black") {
        prize <- 20
      } else if (bet == "odd") {
        prize <- 15
      } else if (bet == "even") {
        prize <- 15
      } else if (bet == "first") {
        prize <- 50
      } else if (bet == "second") {
        prize <- 50
      } else if (bet == "third") {
        prize <- 50
      }
    }
  # add prize from this bet to the total prize
  total_prize <- c(total_prize, prize)
  }
  
  total_prize
}
```

Running the function:

```{r}
set.seed(385)
roulette_loop(many_bets = c("red"))
roulette_loop(many_bets = c("red", "black", "low", "high"))
```

```{r}
long_vec <- rep(c("red", "black", "low", "high", "second", "first", "third",
                  "odd", "even"), 10000)
system.time(roulette_loop(many_bets = long_vec))
```

- Is it taking forever to `knit` the file? If so, add `cache=TRUE` in the chunk above (that is one heavy chunk!)

```{r eval=FALSE}
# write your roulette_vec() function
roulette_vec <- function(many_bets) {
  win_lose_random <- sample(x=c(TRUE,FALSE), size = length(many_bets), replace = TRUE)

  many_bets[many_bets == 'low'] <- 10
  many_bets[many_bets == 'high'] <- 10
  many_bets[many_bets == 'red'] <- 20
  many_bets[many_bets == 'black'] <- 20
  many_bets[many_bets == 'odd'] <- 15
  many_bets[many_bets == 'even'] <- 15
  many_bets[many_bets == 'first'] <- 50
  many_bets[many_bets == 'second'] <- 50
  many_bets[many_bets == 'third'] <- 50
  
  as.numeric(many_bets) * win_lose_random
  
  }
# test your code
set.seed(385)
roulette_vec(many_bets = c("red"))
roulette_vec(many_bets = c("red", "black", "low", "high"))
```

Make sure to check your function using the code segment above. Then, test whether your code is faster using the code segment below:

```{r eval=FALSE}
long_vec <- rep(c("red", "black", "low", "high", "second", "first", "third",
                  "odd", "even"), 10000)
system.time(roulette_vec(many_bets = long_vec))
```

***

## Exercise 2: More Graphs!

Let's learn to graph a few more things! You might need to Google a few things, but that's a part of this course. We learn how to learn coding (with minimal guidance)!

- **(a)** (5 points) **Add a line at x = mean of `Sepal.Length`** to the following histogram

```{r fig.align='center'}
hist(x = iris$Sepal.Length, main = "Histogram of Sepal Length", xlab = "Sepal Length",
     breaks = 15, border = "dodgerblue", probability = TRUE,
     ylim = c(0, 0.65), xlim = c(4, 8.2))
box()
grid()
```

- **(b)** (5 points) Plot a **histogram of the eruption time** of the Old Faith Geyser using the `faithful` dataset, and **add a density estimate curve** to the plot.
  - *Hint*: look up `density()`
  - Use the following plot to check your answer.
  
```{r fig.align='center'}
# Plot the histogram with the density estimate curve
hist(x=faithful$eruptions, xlab = "Eruption Time (mins)", probability = TRUE, main = "Histogram of Old Faitful Geyers Eruption Time", breaks = 15, border = "dodgerblue")
     lines(density(faithful$eruptions), col = 'orange')
```

- **(b)** (5 points) Find a way to plot the following plots as a 1x2 matrix (shown below):
  - Histogram of waiting time from the `faithful` dataset.
  - Scatterplot of `eruptions` vs. `waiting`.
  
```{r fig.align='center'}
# plot the 2 plots
par(mfrow = c(1,2)) 
hist(x = faithful$waiting, xlab = "Waiting Time (mins)", probability = TRUE, main = "Histogram of Waiting Times", breaks = 15, border = "dodgerblue")
plot(x= faithful$waiting, y = faithful$eruptions, col = "dodgerblue", pch = 19)

```


